---
title: "Mass Incarceration"
author: "Abdirahim M, Matthew S, Confidence, Aminata"
date: "2/19/2021"
output: html_document
---

```{r setup, include=FALSE}
source("compile.r")
library(knitr)
library(kableExtra)
```

## Introduction:

*Mass incarceration is an astronomical problem in the United States. The U.S. leads with the highest incarceration rate of any country in the world. Not really something that should be celebrated. Research has shown that it is mainly African-Americans that inhabit these prisons/jails. Is there a societal problem? Can we expose this great injustice to the individuals that, one could say, built this country? Let’s find out!*


*Here are our FIVE variables/ Summary:*

-  According to the data set the race with the highest crime rate in the United States is *`r race_with_the_highest_crime_rate_in_2019`*. Which wasn't surprising to us because the sheer difference of population size between white people and every other race in the United States. We also found out that the crime that was done the most frequently by white people in 2019 was `r highest_crime_type_for_white_in_2019`. This was surprising seeing that in 2013, the crime that was most frequently done by white people was *`r highest_crime_type_for_white_in_2013`*. After some more calculations, We learned that the crimes that  African Americans were mostly arrested for in 2013 was *`r highest_crime_type_for_African_Americans_in_2013`*, which made it even less surprising when we found out that crimes African Americans were arrested for six years later was *`r highest_crime_type_for_African_Americans_in_2019`* because it was very stereotypical. Looking at our data from the years 2013, 2015, 2017, and 2019 we found our that the average number of arrests was *`r average_criminal_total`* throughout the United States.





# Table:

```{r echo=FALSE}
kable(Table[1:4,], caption = "Arrest Info")
```

# Plots
```{r echo=FALSE}
plot(Plot_2019)
```
