---
title: "Mass Incarceration"
author: "Abdirahim M, Matthew S, Confidence, Aminata"
date: "2/19/2021"
output: html_document
---

```{r setup, include=FALSE}
source("compile.r")
library(knitr)
library(kableExtra)
```

## Introduction:

*Mass incarceration is an astronomical problem in the United States. The U.S. leads with the highest incarceration rate of any country in the world. Not really something that should be celebrated.Our data looks at arrests by race for the years 2013,2015,2017, and 2019. The data looks at the total number of arrests along with the number of crimes people are arrested for such as drug crimes, property crimes, and many others. Our data allows users to see what crimes yield the most arrests by race. This data can help use figure out how to reduce arrests by focusing on the crime types that have the biggest impact on these numbers. 

*Here are our FIVE variables/ Summary:*

-  According to the data set the race with the highest crime rate in the United States is *`r race_with_the_highest_crime_rate_in_2019`*. Which wasn't surprising to us because the sheer difference of population size between white people and every other race in the United States. We also found out that the crime that was done the most frequently by white people in 2019 was `r highest_crime_type_for_white_in_2019`. This was surprising seeing that in 2013, the crime that was most frequently done by white people was *`r highest_crime_type_for_white_in_2013`*. After some more calculations, We learned that the crimes that  African Americans were mostly arrested for in 2013 was *`r highest_crime_type_for_African_Americans_in_2013`*, which made it even less surprising when we found out that crimes African Americans were arrested for six years later was *`r highest_crime_type_for_African_Americans_in_2019`* because it was very stereotypical. Looking at our data from the years 2013, 2015, 2017, and 2019 we found our that the average number of arrests was *`r average_criminal_total`* throughout the United States.

Below you can see our table. Some pertinent observations about the table are that arrests are decreasing in the United States. Additionally, we can see that most of the arrests are of whites and blacks  and there are signifcantly more arrests of white people, which is because of population size. This snapshot allows us to see that over half of the arrets for every year are of white people. Also, in 2013 the arrests were very low for Hawaiin pacific islander compared to other years. 

The purpose of the first chart is to show that arrests have been declining from 2013 to 2019. It reveals  that in 2013,2015,2017, and 2019 there was a decline in every year. We included this because it shows a steady decline in arrests which shows that we are heading in the righ direction in fighting mass incarceration, but still more work needs to be done.  I think that this is good news. There has been a lot more awareness around Mass incarceration and this has helped to reduce arrests arrests. ALso, some laws concerning drugs have been changed to help reduce the total number of arrests. 

The purpose of the second chart is to show all the different crimes people are arrested for. This chart allows us to quickly see the crimes that are causing the moost arrests. For instance, it reveals that drug abuse violations, other assaults,property crime, along with a category titled all other offense, which we could not find more information on were the crimes that had the most arrests. We included this becasue we think it is valuable to know what types of crimes are filling up our prison system so that we can create public policy that will help to reduce the number of crimes for this crime type. 




# Table:

```{r echo=FALSE}
kable(Table[1:4,], caption = "Arrest Info")
```

# Plots
```{r echo=FALSE}
plot(Plot_2019)
```
